---
title: "Loading and preprocessing data in DigiRhythm"
author: "Hassan Roland Nasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Loading and preprocessing data in DigiRhythm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>", fig.width = 8
)
```

This Vignette will help you making sure that the data you provide to DigiRhythm follows the library's guideline. We will also make tackle few useful utility functions that might be helpful, especially:

- import_raw_data_activity: a wrapper function that takes data where date and time are in two separate columns and transform a dataset to a digiRhythm friendly data set.
- resample_dgm: this function increases the sampling period of a digiRhythm friendly dataset. Usually, data are acquired in terms with sampling period of seconds or minutes while most of the DigiRhythm functions would be preferably run with a sampling period of 15 minutes.
- is_dgm_friendly: this function will investigate whether a dataset is digiRhythm friendly or not and will output a verbose explaining why.

# Loading data from a sample CSV file
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(digiRhythm)

#A sample dataset could be found here
url <- 'https://raw.githubusercontent.com/nasserdr/digiRhythm_sample_datasets/main/516b_2.csv'
destination <- file.path(tempdir(), '516b_2.csv')
download.file(url, destfile = destination)

# system(paste("head -n 15",  filename)) #Run it only on linux
# IceTag ID:,,50001962,,,,
# Site ID:,,n/a,,,,
# Animal ID:,,n/a,,,,
# First Record:,,30.04.2020,11:54:20,,,
# Last Record:,,15.06.2020,11:06:55,,,
# File Time Zone:,,W. Europe Standard Time,,,,
# 
# Date,Time,Motion Index,Standing,Lying,Steps,Lying Bouts
# 30.04.2020,11:54:20,0,0:00.0,0:40.0,0,0
# 30.04.2020,11:55:00,0,0:00.0,1:00.0,0,0
# 30.04.2020,11:56:00,0,0:00.0,1:00.0,0,0
# 30.04.2020,11:57:00,0,0:00.0,1:00.0,0,0
# 30.04.2020,11:58:00,0,0:00.0,1:00.0,0,0
# 30.04.2020,11:59:00,0,0:00.0,1:00.0,0,0
# 30.04.2020,12:00:00,0,0:00.0,1:00.0,0,0

```

As shown in the previous lines, there are some information that are not needed for working with the data, namely, the first 7 lines. The date and time columns are in two separate columns. In this case, the import_raw_activity_data function is useful. We did not really want to make a universal import function because there might be an unlimited number of cases. However, we only create a function called import_raw_activity_data that solves a couple of most encountered problems, namely the following:

- If the date and time are in two separate columns.
- If the user wants to change the timezone of the data (by using the arguments original_tz and target_tz).
- If the user wants to change directly the sampling period.
- If the user wants to remove the dates that do not contains a whole day recording by using trim_first_day or trim_last_day or trim_middle_days. If one of these arguments is TRUE, then the function will remove the data of the first (or last day, respectively) if they contains data less than 80% of what they are supposed to contains.

For our particular example, as mentionned in the function call below, we would like to skip 7 lines, we need to read 4 columns, we specified the date and time formats, the separator, the original and target time zones, the sampling rate and we want to remove all days that contains less than 80% of data.

```{r Importing}
data <- import_raw_activity_data(destination,
                                   skipLines = 7,
                                   act.cols.names = c("Date", "Time", "Motion Index", 'Steps'),
                                   date_format = "%d.%m.%Y",
                                   time_format = "%H:%M:%S",
                                   sep = ',',
                                   original_tz = 'CET',
                                   target_tz = 'CET',
                                   sampling = 15,
                                   trim_first_day = TRUE,
                                   trim_middle_days = TRUE,
                                   trim_last_day = TRUE,
                                   verbose = TRUE)

```

As Shown, the argument verbose = TRUE outputs some useful information about the data loading process. We particularly mention the following output:
Minimum Required number of samples per day 76. The 76 is obtained by computing 80% of the hypothetical data samples. In fact, if the sampling period is 15 min, then we are supposed to have 96 samples per day. 80% of the 96 samples is equal to 76, therefore, days with less than 96 data points are removed.

# Checking if the data is digiRhythm friendly
Imposing a specific format on the dataset is a design choice. We wanted to avoid future
possible errors that might happen when user use the library. A digiRhythm friendly dataset
has the following properties:
- It has at least two columns.
- The first column should be of a POSIX format (datetime).
- All other columns should be numeric.
- The dataset should contains data for 7 days at least (7 days are required to compute the DFC).

The function [dgm_friendly()] returns a boolean. If verbose is set to TRUE, it will output
details about why or why not the data is not considered digiRhythm friendly.

```{r dgm_friendly}
is_dgm_friendly(data, verbose = TRUE)
```

# Removing outliers

Sometimes, we may have outliers and want to remove them. Although the outlier removal
routine is not complicated, but providing an off-the-shelve function for DGM friendly data
is useful and might save time. The [remove_activity_outliers()] (available inside utils.R) will remove the outliers in all the non-POSIX columns. All data points that lie out of the 2.5 and 97.5 percentiles are considered outliers. 
```{r outliers}
data_without_outliers <- remove_activity_outliers(data)
head(data_without_outliers)
```


# Resampling data

Often, data are acquired with milliseconds or seconds sampling periods. However, transforming data to higher sampling periods is recommended for some algorithms like the degree of functional coupling. The library offers resampling function as it is useful in this case. However, here are two considerations here:

- The sampling period can be increased but not decreased. For instance, we can transform data from 1 min sampling period to 15 minutes sampling period.
- The resampling concerns only activity variables where it makes sense to make an additive resampling. For instance, if we are measuring the number of steps every 1 min and we want to resample it to 15 minutes, we can simply add all the steps during these 15 minutes. For other types of variables, this might not be the case. For example, if we are measuring the speed or the acceleration, addition would not probably make sense, and a further transformation is required (in this case, it would be dividing by 15 to obtain the average speed or acceleration). In other case where the variable reflect a categorical entity that could be transformed into a dummy variable, we argue that resampling would not at all make sense.

```{r resampling}
resampled_data <- resample_dgm(data, new_sampling = 15)
head(resampled_data)
```

# Periodicity of a dataset

Users can have access to the periodicity of a dataset using the dgm_periodicity function as shown below:

```{r periodicity}
s <- dgm_periodicity(data)
```
